---
author: "Daniel Weinberger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
params:
  sensitivity: TRUE
  crossval: FALSE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 3,
  fig.width = 5,
  fig.align = "center", 
  dpi=300, 
	out.width="600px"
)
```

---

```{r setup_packages, include=FALSE, echo=TRUE}
#Install the package
#library(devtools)
#devtools::install_github('https://github.com/weinbergerlab/InterventionEvaluatR') 
library(knitr)
library(plyr)
library(InterventionEvaluatR)
library(rjags)
library(coda)
library(HDInterval)
library(lubridate)
library(pbapply)
library(parallel)
library(plyr)
```

---
title: "Estimated change associated with the introduction of vaccine in South Africa"
---

---
## Important note
This Rmd file contains analyses of the data from S Africa. It uses a slightly different model (Poisson) than the one described in Kleynhans et al. To see those results, see the synthetic_control_run.R file

```{r viewdata}
   sa1<-read.csv('../Data/RSA.csv')

   #Filter covariates that have no variance
   sa.covars<-sa1[, -which(names(sa1) %in% c('Pneum','date','age')  )]
   sa.covars$index<-1:nrow(sa.covars)
   sa1.spl<-split(sa.covars, sa1$age)
   variance.vars<-lapply(sa1.spl, function(x) apply(x[1:123,], 2, var)  ) #look in pre-vax period
      for(i in 1:length(sa1.spl)){
     sa1.spl[[i]]<-sa1.spl[[i]][,-which(variance.vars[[i]]==0)]
   }
   sa.covars<-rbind.fill(sa1.spl) #cobine back together
   sa.covars<-sa.covars[order(sa.covars$index),]
   sa.covars$index<-NULL
   sa2<-cbind.data.frame(sa1[, which(names(sa1) %in% c('Pneum','date','age')  )], sa.covars)
   sa2$date<-as.Date(sa1$date, '%Y-%m-%d' ) 
   
   exclude_covar <- c('AllRes', 	'PI', 	'POth', 	'AllDiar', 	 	'Menin', 	'BactMenin', 	'Bact', 	'OtMed', 	'Mast',	'SeptArthr', "denom", "A20_B99_excl_bac", "A16", "A17", "A18", "A19", "R00_R09", "R10_R19", "R20_R39", "R40_R49", "R50_R69", "R70_R94", "R95_R99", "D50_D89", "R00_R99")      

   sa2<-sa2[,-which(names(sa2) %in% exclude_covar)]
   
 
```

#Plot time series
```{r, fig.width=8}
unique(sa2$age)
age_subset<-sa2[sa2$age=='40-64 years',]
par(mfrow=c(1,2))
plot(age_subset$Pneum, type='l', bty='l')
abline(v=123, col='gray', lty=2)
plot(age_subset$R00_R99, type='l', bty='l')
abline(v=123, col='gray', lty=2)


matplot(scale(log(age_subset[,4:20]+0.5)), type='l', bty='l')
abline(v=123, col='gray', lty=2)


plot(age_subset$Pneum)

prePneum<-age_subset$Pneum
prePneum[124:length(prePneum)] <-NA
ts1<-log(age_subset$Pneum)
plot(ts1)
lag12<-c(rep(NA,12),ts1[1:(length(ts1)-12)])
lag1<- c(NA,ts1[1:(length(ts1)-1)])
log.a16<-log(age_subset$A16_A19+1)
log.r00<-log(age_subset$R00_R99+1)
logj00<-log(age_subset$J00_J99_excl_PI_bron+1)
log.b20<-log(age_subset$B20_B24+1)

#Lag1 and lag 12 + regressor
mod1<-glm(prePneum~ lag12+lag1 +log.a16 , family='poisson')
summary(mod1)

#lag12 + regressor
mod2<-glm(prePneum~ lag12+log.a16 , family='poisson')
summary(mod2)

#Monthly dummy+ regressor
mod.ds<-cbind.data.frame(prePneum,'month'=as.factor(month(age_subset$date)), 'log.a16'=log.a16,log.r00, logj00,log.b20)
mod3<-glm(prePneum~ month+  log.b20 +log.a16, family='quasipoisson', data=mod.ds)
summary(mod3)
pred3<-predict(mod3, type='response', newdata=mod.ds)
plot(age_subset$date,age_subset$Pneum/pred3)
abline(v=analysis$intervention_date)
abline(h=1)

#GLARMA
library(glarma)
Y<-prePneum[1:123]
X<- model.matrix(~month+  log.b20  +log.a16, data=mod.ds[1:123,])
glarmamod <- glarma(Y, X, phiLags = c(1), type = "Poi", method = "NR" , residuals = "Pearson", maxit = 100, grad = 1e-6)
 summary(glarmamod)

 
#lag1+lag12+ offset
mod4<-glm(prePneum~  lag12+lag1 + offset(log.r00) , family='poisson')
summary(mod4) 
 
#pre period is 1:123
plot(age_subset$date,age_subset$Pneum/age_subset$R00_R99)
abline(v=analysis$intervention_date)
```


# Intervention EvaluatR Package

## Set parameters for analysis

Here we need to set a few parameters. We Use the evaluatr.init() function to specify the name of the dataset, the date at which the vaccine is introduced, the date at which we want to begin evaluating the vaccine (typically 1-2 year after vaccine introduction). We also provide some information on the dataset, sch as whether the data are monthly or quarterly (n_seasons), the variable names for the grouping variable, the date variable, the outcome variable, and the denominator variable (if any).

```{r setup_data, echo=TRUE}

analysis <- evaluatr.init(
  country = "SAfrica", data = sa2,
  post_period_start = "2009-04-01", 
  eval_period_start = "2010-04-01", 
  eval_period_end = "2016-12-01", 
  n_seasons = 12, #This is monthly data, so select 12
  year_def = "cal_year", # we are in southern hemisphere, so aggregate results by calendar year (Jan-Dec)
  group_name = "age",  #Strata categry name
  date_name = "date", #Date variable name
  outcome_name = "Pneum", #Outcome variable name
  denom_name = "denom", #Denominator variable name
  set.burnN=5000,
  set.sampleN=10000
)
set.seed(1)
```



## Run a simple analysis controlling for 1 control variable at a time

Before getting into more complicated analyses, we will first try to fit a simple Poisson regression model (with overdispersion) where we adjust for seasonality and 1 control variable at a time. this allows us to see how the use of different controls influences the results

The results are ordered by goodness of fit (based on AIC scores), with best fitting covariates on top.
```{r univariate, fig.width=3, fig.height=5}
 glmer_results= evaluatr.univariate(analysis)
 lapply(glmer_results,evaluatr.univariate.plot)
```

Compare a simple Poisson model with a simple log-Gaussian model
```{r}


ds1<-analysis$input_data[analysis$input_data$age=='80 years and above',]
covars<-ds1[,4:ncol(ds1)]
covars<-apply(covars,2, function(x) scale(log(x+0.5))   )
ds2<-cbind.data.frame(ds1[,1:3],covars)
ds2$month<-month(ds2$date)
ds2$outcome<- ds2$Pneum
dates<-unique(ds2$date)
ds2$outcome.pre<-ds2$outcome
ds2$outcome.pre[dates > analysis$pre_period_end] <-NA
mod1<- glm(outcome.pre ~ as.factor(month)+  R00_R99 +J00_J99_excl_PI_bron  +I00_I99, family='poisson', data=ds2)
summary(mod1)
pred1<-predict(mod1, type='response', newdata=ds2)
sum.post.obs<- sum(ds2$outcome[dates > analysis$pre_period_end])
sum.post.pred<- sum(pred1[dates > analysis$pre_period_end])
rr1<-sum.post.obs/sum.post.pred



ds2$log.outcome.pre<-log(ds2$outcome.pre+0.5)
mod2<- glm(log.outcome.pre ~as.factor(month) +   R00_R99 +J00_J99_excl_PI_bron  +I00_I99 , family='gaussian', data=ds2)
summary(mod2)
pred2<-exp(predict(mod2, type='response', newdata=ds2))
sum.post.obs<- sum(ds2$outcome[dates > analysis$pre_period_end])
sum.post.pred2<- sum(pred2[dates > analysis$pre_period_end])
rr2<-sum.post.obs/sum.post.pred2

rr1
rr2


matplot(covars[,c('R00_R99','J00_J99_excl_PI_bron','I00_I99')], type='l')
```


## Run the main analysis
Save the results in object 'impact_results'
```{r main analysis, include = FALSE}
impact_results = evaluatr.impact(analysis)
```

##Run sensitivity analyses
Sequentially drop the top 1,2, or 3 variables in synthetic controls analysis
```{r sensitivity_analyses, include = FALSE}
if (params$sensitivity) {
  sensitivity_results <- evaluatr.sensitivity(analysis)
}
```


#`r params$country` Results


```{r sparse}
if (!is.null(names(analysis$sparse_groups[analysis$sparse_groups])) && length(names(analysis$sparse_groups[analysis$sparse_groups])) != 0) {
  kable(data.frame("Sparse Groups" = names(analysis$sparse_groups[analysis$sparse_groups]), check.names = FALSE), align = "c")
}
```

## Check model convergence
These models are fit using Markov Chain Monte Carlo (MCMC). It is important to evaluate the convergence of the model. A quick way to check this is to evaluate the trace plots for rate ratio estimates from the synthetic controls model (or for the other model variants). We use Geweke's diagnostic, which tests whether the mean estimate for the rate ratio in the first 10% of the iterations is equal to the mean estimate for the rate ration in the last 50% of the iterations. If the model has not converged, you might need to add more iterations or a longer burn in period
```{r}
all.traces<-sapply(impact_results,'[[', 'rr_iter')
cats<-dimnames(all.traces[[1]])[[1]]
all.converge.status<-sapply(impact_results,'[[', 'converge')
for(j in c('full','time','pca')){
for(i in 1: nrow(all.traces[[1]])){
   plot(all.traces[[j]][i,], type='l', main=paste0(j,' ',cats[i],' ' ,all.converge.status[[j]][i,2] ), bty='l', ylim=c(0.2,2))
}
}
```

##combine estimates

```{r Comparison of estimates from different models}
if (params$crossval) {
  kable(cbind.data.frame(crossval_results$rr_mean_stack_intervals, impact_results$full$rr_mean_intervals, impact_results$time$rr_mean_intervals, impact_results$time_no_offset$rr_mean_intervals, impact_results$its$rr_mean_intervals, impact_results$pca$rr_mean_intervals), align = "c")
} else {
  kable(cbind.data.frame(impact_results$best$rr_mean_intervals, impact_results$full$rr_mean_intervals, impact_results$time$rr_mean_intervals, impact_results$time_no_offset$rr_mean_intervals, impact_results$its$rr_mean_intervals, impact_results$pca$rr_mean_intervals), align = "c")
}
```

##Plot of Rate ratios, with size proportional to cross validation weights
```{r mainplot1, echo=FALSE}
plots <- evaluatr.plots(analysis)
plots$summary
```


##Number of variables selected in SC analysis
```{r modelsize}
kable(analysis$model_size, col.names = c("Model Size"))
```

##Inclusion Probabilities
```{r incl, include = FALSE}
incl_probs <- NULL
for (group in analysis$groups) {
  incl_prob <- impact_results$full$groups[[group]]$inclusion_probs[-c(1:(analysis$n_seasons - 1)), ]
  incl_prob <- incl_prob[order(-incl_prob$inclusion_probs), ]
  incl_prob <- incl_prob[c(1:3), ]
  incl_prob2 <- incl_prob[, 2]
  incl_prob_names <- incl_prob[, 1]
  incl_prob3 <- data.frame("Group" = group, "Greatest Inclusion Variable" = incl_prob_names[1], "Greatest Inclusion Probability" = incl_prob2[1], "Second Greatest Inclusion Variable" = incl_prob_names[2], "Second Greatest Inclusion Probability" = incl_prob2[2], "Third Greatest Inclusion Variable" = incl_prob_names[3], "Third Greatest Inclusion Probability" = incl_prob2[3], check.names = FALSE)
  incl_probs <- rbind(incl_probs, incl_prob3)
}
rownames(incl_probs) <- NULL
```

```{r incl_table}
kable(incl_probs, align = "c")
```

## Weight Sensitivity Analysis
```{r sensitivity}
if (exists("sensitivity_results")) {
  kable(sensitivity_results$sensitivity_table_intervals, align = "c")
}
```


## Plot Observed vs expected monthly time series
```{r plots, results = 'asis', plot.width=5, plot.height=12}
for (group in names(plots$groups)) {
      par(mfrow=c(4,1))
      print(plots$groups[[group]]$pred_full )
     # print(plots$groups[[group]]$pred_best )
      #print(plots$groups[[group]]$pred_time )
      #print(plots$groups[[group]]$pred_pca )
}
```

## Plot Observed vs expected yearly time series
```{r plots2, results = 'asis', plot.width=5, plot.height=12}
for (group in names(plots$groups)) {
      par(mfrow=c(4,1))
      print(plots$groups[[group]]$pred_full_agg )
      print(plots$groups[[group]]$pred_best_agg )
      print(plots$groups[[group]]$pred_time_agg )
      print(plots$groups[[group]]$pred_pca_agg )
}
```

## Plot cumulative cases prevented
Estimated using the 'best' model, between SC and STL+PCA

```{r plots3, results = 'asis', plot.width=5, plot.height=12}
for (group in names(plots$groups)) {
      par(mfrow=c(4,1))
      print(plots$groups[[group]]$cumsum_prevented )
}
```

## Print results
```{r save_results, echo=FALSE}
output_file <- "Results" # Directory where results will be saved.
output_file <- paste0(output_file, "_", analysis$country, "_", format(Sys.time(), "%Y-%m-%d-%H%M%S"), ".Rds")
evaluatr.save(analysis, output_file)
```


Try AR1 horseshoe model. The AR1 model seems to have an issue where the AR1 component competes with the regression piece and displaces the regresison effect. This is undesirable here because we want the regressors to explain the trends, not AR1
```{r}
call.jags.mod<-function(ds.in){  
  exclude.cols<-c(analysis$group_name, analysis$outcome_name, analysis$date_name)
  x<-ds.in[,-which(names(ds.in) %in% exclude.cols)]
    
  #Filter columsn with 0 variations in the covariate in the pre-vax period
  x.var<-apply(x,2, function(xx) var(xx[ds.in[,analysis$date_name]<analysis$post_period[1]] ))  
  x.var[is.na(x.var)]<-0
  x<-x[,x.var>0] 
  
  x.scale<-apply(x,2, function(z) scale(log(z+0.5))) 
  y<-ds.in[,analysis$outcome_name] 
  ds2<-cbind.data.frame(y, x.scale) 
  names(ds2)<-c(analysis$outcome_name,names(x)) 
  ds2.pre<-ds2[ds.in[,analysis$date_name]<analysis$post_period[1] ,] 
  ds3<-ds2 
  ds3[,analysis$outcome_name][ds.in[,analysis$date_name]>=analysis$post_period[1]]<-NA 
  months<-month(ds.in[,analysis$date_name]) 
  month.mat<-dummies::dummy(months) 
  month.mat<-month.mat[,-1]
  month.mat<-cbind(rep(1,nrow(month.mat)), month.mat) #add intercept
  month.mat.pre<- month.mat[ds.in[,analysis$date_name]<analysis$post_period[1] ,]
  
  mod.txt<- source('../jags/Non_spatial-Non_lagged-iid.txt')
  mod1<-nonspace_nonlag(burnin=70000, 
                        samples=40000,
                        thin=1,
                        chains=1,
                        regularize=TRUE, 
                        dic=FALSE,
                        n_full=length(ds3[,analysis$outcome_name]),
                        n_modeling=nrow(ds2.pre),
                        y_modeling=ds3[,analysis$outcome_name],
                        offset=rep(0, length(y)) ,
                        z=month.mat,  #parameters not being shrunk
                        x=ds3[,-1]
                        )
  
  posterior_samples.all<-do.call(rbind,mod1[[1]])
  post_means<-apply(posterior_samples.all, 2, median)
  sample.labs<-names(post_means)
  ci<-t(hdi(posterior_samples.all, credMass = 0.95))
  row.names(ci)<-sample.labs
  names(post_means)<-sample.labs
  post.combo<-cbind(post_means,ci)
  
  post.combo.y<-post.combo[grep('Y_pred', dimnames(post.combo)[[1]]),]
  post.combo.beta<-cbind.data.frame(post.combo[grep('beta', dimnames(post.combo)[[1]]),],dimnames(x)[[2]])
  
  post.combo.phi<- post.combo[grep('phi', dimnames(post.combo)[[1]]),]

  # plot(y=1:nrow(post.combo.beta), x=post.combo.beta[,'post_means'], xlim=range(post.combo.beta[,1:3]), bty='l')
  # arrows(y0=1:nrow(post.combo.beta) ,x0=post.combo.beta[,2], x1=post.combo.beta[,3], length=0)
  # abline(v=0, col='gray', lty=2)
  # text(y=1:nrow(post.combo.beta), x=post.combo.beta[,'post_means']+0.1,dimnames(x)[[2]] , col='gray')
  
  #plot for post-vax period
  # matplot(post.combo.y, type='l')
  #  points(ds1$sALRI[(nrow(ds2.pre)+1):nrow(ds1)])
  # 
    log.rr.pointwise<- log((ds.in[(nrow(ds2.pre)+1):nrow(ds.in),analysis$outcome_name]+0.5)/(post.combo.y+0.5))
   # matplot(log.rr, type='l', col='gray', lty=c(2,1,2))
   # abline(h=0, col='red')
   
   length.rollout<-round(as.numeric((analysis$eval_period[1]-analysis$post_period[1]  )/30.3))
   post.samples<-posterior_samples.all[,-c(1:length.rollout)]
   post.samples.y<- post.samples[,grep('Y_pred', dimnames(post.samples)[[2]])]

   post.samples.sum<-apply(post.samples,1,sum)
   obs.post.sum<- sum(ds2[,analysis$outcome_name][ds.in[,analysis$date_name]>=analysis$post_period[1]][-c(1:length.rollout)])
   rr.agg<-obs.post.sum/post.samples.sum
   rr.q<-quantile(rr.agg, probs=c(0.025, 0.5, 0.975))
   output.list<-list('rr.samples'=rr.agg,'rr.q'=rr.q,'log.rr.pointwise.q'=log.rr.pointwise,'betas'=post.combo.beta,'rr.agg.iter'=rr.agg,'phi'=post.combo.phi)
}
```

```{r}
n_cores<-7
grp.keep<-c("Less than one month", "1-11 months", "1-4 years", "5-7 years", "8-18 years","19-39 years","65-79 years" )

ds.age.spl<-split(sa2, sa2$age)
ds.age.spl<-ds.age.spl[grp.keep]

  cl <- makeCluster(n_cores)
  clusterEvalQ(cl, {
    library(lubridate, quietly = TRUE)
    library(HDInterval, quietly = TRUE)
  })
  clusterExport(cl, c('call.jags.mod','analysis'), environment())
  
  mod1<-pblapply(cl = cl,X=ds.age.spl,FUN=call.jags.mod)
stopCluster(cl)

rr.samples<-sapply(mod1, '[[', 'rr.samples',simplify='array')
phi<-sapply(mod1, '[[', 'phi',simplify='array')

```




```{r}
rr.summary<-t(apply(rr.samples[30000:40000,],2,quantile, probs=c(0.025,0.5,0.975)))
rr.summary
```

```{r}
matplot(phi[,,2 ], type='l')
abline(h=0)

```

Length of burn in required varies by dataset. many are fine with 15-20000 (+10K original burn). By 0K (+10K original burn) all look good. So could prob get away with a burn of 60K + 10K samples (70K)?
```{r}
apply(rr.samples,2, function(x) plot(x,type='l'))
```


```{r}
betas<-sapply(mod1, '[[', 'betas',simplify=F)
betas<-lapply(betas, function(x) x[order(-x$post_means),])
names(betas)
betas[[2 ]]
```



## PCA regression+ AR1
```{r}
call.jags.mod.pc<-function(ds.in){  
  exclude.cols<-c(analysis$group_name, analysis$outcome_name, analysis$date_name)
  x<-ds.in[,-which(names(ds.in) %in% exclude.cols)]
    
  #Filter columsn with 0 variations in the covariate in the pre-vax period
  x.var<-apply(x,2, function(xx) var(xx[ds.in[,analysis$date_name]<analysis$post_period[1]] ))  
  x.var[is.na(x.var)]<-0
  x<-x[,x.var>0] 
  
  x.scale<-apply(x,2, function(z) scale(log(z+0.5))) 
  pca1<- prcomp(x.scale, center = TRUE,scale. = TRUE)
  pcs<-pca1$x
  x.pcs<-pcs[,1:2]
  
  y<-ds.in[,analysis$outcome_name] 
  ds2<-cbind.data.frame(y, x.pcs) 
  names(ds2)<-c(analysis$outcome_name,names(x.pcs)) 
  ds2.pre<-ds2[ds.in[,analysis$date_name]<analysis$post_period[1] ,] 
  ds3<-ds2 
  ds3[,analysis$outcome_name][ds.in[,analysis$date_name]>=analysis$post_period[1]]<-NA 
  months<-month(ds.in[,analysis$date_name]) 
  month.mat<-dummies::dummy(months) 
  month.mat<-month.mat[,-1]
  month.mat<-cbind(rep(1,nrow(month.mat)), month.mat) #add intercept
  month.mat.pre<- month.mat[ds.in[,analysis$date_name]<analysis$post_period[1] ,]
  
  mod.txt<- source('../jags/Non_spatial-Non_lagged-iid.txt')
  mod1<-nonspace_nonlag(burnin=70000,  #70K for AR1
                        samples=40000, #40K for AR1
                        thin=1,
                        chains=1,
                        regularize=FALSE, 
                        dic=FALSE,
                        n_full=length(ds3[,analysis$outcome_name]),
                        n_modeling=nrow(ds2.pre),
                        y_modeling=ds3[,analysis$outcome_name],
                        offset=rep(0, length(y)) ,
                        z=month.mat,  #parameters not being shrunk
                        x=ds3[,-1]
                        )
  
  posterior_samples.all<-do.call(rbind,mod1[[1]])
  post_means<-apply(posterior_samples.all, 2, median)
  sample.labs<-names(post_means)
  ci<-t(hdi(posterior_samples.all, credMass = 0.95))
  row.names(ci)<-sample.labs
  names(post_means)<-sample.labs
  post.combo<-cbind(post_means,ci)
  
  post.combo.y<-post.combo[grep('Y_pred', dimnames(post.combo)[[1]]),]
  post.combo.beta<-cbind.data.frame(post.combo[grep('beta', dimnames(post.combo)[[1]]),],dimnames(x.pcs)[[2]])
  
  post.combo.phi<- post.combo[grep('phi', dimnames(post.combo)[[1]]),]

  # plot(y=1:nrow(post.combo.beta), x=post.combo.beta[,'post_means'], xlim=range(post.combo.beta[,1:3]), bty='l')
  # arrows(y0=1:nrow(post.combo.beta) ,x0=post.combo.beta[,2], x1=post.combo.beta[,3], length=0)
  # abline(v=0, col='gray', lty=2)
  # text(y=1:nrow(post.combo.beta), x=post.combo.beta[,'post_means']+0.1,dimnames(x)[[2]] , col='gray')
  
  #plot for post-vax period
  # matplot(post.combo.y, type='l')
  #  points(ds1$sALRI[(nrow(ds2.pre)+1):nrow(ds1)])
  # 
    log.rr.pointwise<- log((ds.in[(nrow(ds2.pre)+1):nrow(ds.in),analysis$outcome_name]+0.5)/(post.combo.y+0.5))
   # matplot(log.rr, type='l', col='gray', lty=c(2,1,2))
   # abline(h=0, col='red')
   
   length.rollout<-round(as.numeric((analysis$eval_period[1]-analysis$post_period[1]  )/30.3))
   post.samples<-posterior_samples.all[,-c(1:length.rollout)]
   post.samples.y<- post.samples[,grep('Y_pred', dimnames(post.samples)[[2]])]

   post.samples.sum<-apply(post.samples,1,sum)
   obs.post.sum<- sum(ds2[,analysis$outcome_name][ds.in[,analysis$date_name]>=analysis$post_period[1]][-c(1:length.rollout)])
   rr.agg<-obs.post.sum/post.samples.sum
   rr.q<-quantile(rr.agg, probs=c(0.025, 0.5, 0.975))
   output.list<-list('rr.samples'=rr.agg,'rr.q'=rr.q,'log.rr.pointwise.q'=log.rr.pointwise,'betas'=post.combo.beta,'rr.agg.iter'=rr.agg,'phi'=post.combo.phi)
}

```

```{r}
n_cores<-7
grp.keep<-c("Less than one month", "1-11 months", "1-4 years", "5-7 years", "8-18 years","19-39 years","65-79 years" )

ds.age.spl<-split(sa2, sa2$age)
ds.age.spl<-ds.age.spl[grp.keep]

  cl <- makeCluster(n_cores)
  clusterEvalQ(cl, {
    library(lubridate, quietly = TRUE)
    library(HDInterval, quietly = TRUE)
  })
  clusterExport(cl, c('call.jags.mod.pc','analysis'), environment())
  
  mod1.pc<-pblapply(cl = cl,X=ds.age.spl,FUN=call.jags.mod.pc)
stopCluster(cl)

rr.samples<-sapply(mod1.pc, '[[', 'rr.samples',simplify='array')
phi<-sapply(mod1.pc, '[[', 'phi',simplify='array')
```


```{r}
rr.summary<-t(apply(rr.samples,2,quantile, probs=c(0.025,0.5,0.975)))
rr.summary
```

```{r}
apply(phi, 3, matplot, type='l')

```

Length of burn in required varies by dataset. many are fine with 15-20000 (+10K original burn). By 0K (+10K original burn) all look good. So could prob get away with a burn of 60K + 10K samples (70K)?
```{r}
apply(rr.samples[8000:10000,],2, function(x) plot(x,type='l'))
```


```{r}
betas<-sapply(mod1.pc, '[[', 'betas',simplify=F)
betas<-lapply(betas, function(x) x[order(-x$post_means),])
names(betas)
betas
```



Simple GLM model with PC1 and PC2
```{r}
pc.func<-function(age.select){
  x<-sa2[sa2$age==age.select,4:40]
  y<-  sa2[sa2$age==age.select,"Pneum"]
  dates<-unique(sa2$date)
  y.pre<-y
  y.pre[dates>=analysis$intervention_date]<-NA
  #Filter columsn with 0 variations in the covariate in the pre-vax period
  x.var<-apply(x,2, function(xx) var(xx))  
  x.var[is.na(x.var)]<-0
  x<-x[,x.var>0] 
  
  x.scale<-apply(x,2, function(z) scale(log(z+0.5))) 
pca1<- prcomp(x.scale, center = TRUE,scale. = TRUE)

pcs<-pca1$x
rotation<-pca1$rotation

#plot(pcs[,'PC1'], type='l')
#plot(pcs[,'PC2'], type='l')

#library(devtools)
#install_github("vqv/ggbiplot")
library(ggbiplot)
ggbiplot(pca1)
screeplot(pca1)

mod.df<-cbind.data.frame(y.pre, 'month'=as.factor(month(dates)), y,'pc1'=pcs[,'PC1'],'pc2'=pcs[,'PC2'],'pc3'=pcs[,'PC3'],'pc4'=pcs[,'PC4'] )
mod1<- glm(y.pre ~ pc1+pc2 +month, family='quasipoisson', data=mod.df)
print(age.select)
print(summary(mod1))
pred1<-predict(mod1, type='response', newdata=mod.df)

plot(dates,pred1, type='l', main=age.select)
points(dates, y)

plot(dates,y/pred1, type='l', main=age.select)
abline(h=1)
abline(v=analysis$intervention_date)
}

lapply(unique(sa2$age), pc.func)
```


GLARMA model with PC1 and PC2
```{r}
library(glarma)
age.select<-unique(sa2$age)[2]

pc.func<-function(age.select){
  x<-sa2[sa2$age==age.select,4:40]
  y<-  sa2[sa2$age==age.select,"Pneum"]
  dates<-unique(sa2$date)
  y.pre<-y[dates<analysis$intervention_date]
  #Filter columsn with 0 variations in the covariate in the pre-vax period
  x.var<-apply(x,2, function(xx) var(xx))  
  x.var[is.na(x.var)]<-0
  x<-x[,x.var>0] 
  
  x.scale<-apply(x,2, function(z) scale(log(z+0.5))) 
  
pca1<- prcomp(x.scale, center = TRUE,scale. = TRUE)

pcs<-pca1$x
rotation<-pca1$rotation

#plot(pcs[,'PC1'], type='l')
#plot(pcs[,'PC2'], type='l')

#library(devtools)
#install_github("vqv/ggbiplot")
library(ggbiplot)
ggbiplot(pca1)
screeplot(pca1)

mod.df<-cbind.data.frame(y.pre, 'month'=as.factor(month(dates)), y,'pc1'=pcs[,'PC1'],'pc2'=pcs[,'PC2'],'pc3'=pcs[,'PC3'],'pc4'=pcs[,'PC4'] )
#mod1<- glm(y.pre ~ pc1+pc2 +month, family='quasipoisson', data=mod.df)
X<- model.matrix(~ month+  pc1  +pc2, data=mod.df)
X.pre<-X[dates<analysis$intervention_date,]
mod.df.post<-X[dates>=analysis$intervention_date,]
  
mod1 <- glarma(y=y.pre, X=X.pre, phiLags = c(1), type = "Poi", method = "NR" , residuals = "Pearson", maxit = 100, grad = 1e-6)

print(age.select)
print(summary(mod1))
pred1<-forecast(mod1, type='response', newdata=mod.df.post, n.ahead = 2)

plot(dates,pred1, type='l', main=age.select)
points(dates, y)

plot(dates,y/pred1, type='l', main=age.select)
abline(h=1)
abline(v=analysis$intervention_date)
}

lapply(unique(sa2$age), pc.func)
```

INLA AR1
```{r}
#install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
library(INLA)
#age.select<-unique(sa2$age)[3]
inla.ar1.func<-function(age.select){
  x<-sa2[sa2$age==age.select,4:40]
  y<-  sa2[sa2$age==age.select,"Pneum"]
  dates<-unique(sa2$date)
  y.pre<-y
  y.pre[dates>=analysis$intervention_date]<-NA
  #Filter columsn with 0 variations in the covariate in the pre-vax period
  x.var<-apply(x,2, function(xx) var(xx))  
  x.var[is.na(x.var)]<-0
  x<-x[,x.var>0] 
  
  x.scale<-apply(x,2, function(z) scale(log(z+0.5))) 
  
pca1<- prcomp(x.scale, center = TRUE,scale. = TRUE)

pcs<-pca1$x
rotation<-pca1$rotation

mod.df<-cbind.data.frame(y.pre, 'month'=as.factor(month(dates)), y,'pc1'=pcs[,'PC1'],'pc2'=pcs[,'PC2'],'pc3'=pcs[,'PC3'],'pc4'=pcs[,'PC4'] )
#mod1<- glm(y.pre ~ pc1+pc2 +month, family='quasipoisson', data=mod.df)
mod.df$t<-1:nrow(mod.df)
mod.df.pre<-mod.df[dates<analysis$intervention_date,]

inla.mod1 <- inla(y.pre ~ pc1+pc2 +month+ f(t, model = "ar1"),data=mod.df,
    family='poisson',
  control.predictor = list(compute = TRUE, link = 1),
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE, config=TRUE),
                       E= rep(1, nrow(mod.df))
)
posterior.list<-inla.posterior.sample(n=100, inla.mod1)
post.labels<-dimnames(posterior.list[[1]]$latent)[[1]]
posterior.samples<- sapply(posterior.list, '[[', 'latent')
preds.select<-grepl('Predictor',post.labels )
posterior.preds<-exp(posterior.samples[preds.select,]) #lambda
#now take Poisson samples withmean of lambda
posterior.preds.counts<- matrix(rpois(n=length(posterior.preds), lambda=posterior.preds), nrow=nrow(posterior.preds), ncol=ncol(posterior.preds))

posterior.preds.q<-t(apply(posterior.preds.counts,1,quantile, probs=c(0.025, 0.5, 0.975)))
posterior.median<-as.integer(round(t(apply(posterior.preds.counts,1,median))))
ci<- t(hdi(t(posterior.preds.counts), credMass = 0.95))
posterior.pred.hdi<- cbind.data.frame('median'=posterior.median, ci)


   log.rr.pointwise<- apply(posterior.preds.counts, 2, function(x)log( (mod.df$y+1)/ (x+1)) )
   log.rr.pointwise.ci<- t(hdi(t(log.rr.pointwise), credMass = 0.95))
   log.rr.pointwise.median<- apply(log.rr.pointwise.ci,1,median)
   log.rr.pointwise.hdi<-cbind(log.rr.pointwise.median,log.rr.pointwise.ci)
   # matplot(log.rr, type='l', col='gray', lty=c(2,1,2))
   # abline(h=0, col='red')
   
   post.period<- which(dates>= analysis$eval_period[1]  &dates<= analysis$eval_period[2] )
   
   post.samples<-posterior.preds.counts[post.period,]

   post.samples.sum<-apply(post.samples,2,sum)
   obs.post.sum<- sum(mod.df$y[post.period])
   rr.agg<-obs.post.sum/post.samples.sum
   rr.q<-quantile(rr.agg, probs=c(0.025, 0.5, 0.975))
   rr.hdi<-c(rr.q['50%'],hdi(rr.agg, credMass = 0.95))

out.list<- list('posterior.pred.hdi'=posterior.pred.hdi, 'rr.hdi'=rr.hdi,'pca.results'=pca1,'pcs'=pcs,'inla.mod'=inla.mod1,'log.rr.pointwise.hdi'=log.rr.pointwise.hdi,'obs.y'=mod.df$y)

}
inla1<-lapply(unique(sa2$age), inla.ar1.func)

```

RR for evaluation period
```{r}
rr<-cbind.data.frame(unique(sa2$age),round(t(sapply(inla1,'[[', 'rr.hdi')),2))
rr
```

RR for each time points
```{r}
pointwise.log.rr<-(sapply(inla1,'[[', 'log.rr.pointwise.hdi', simplify='array') )
apply(exp(pointwise.log.rr),3, function(x){
  matplot(x ,type='l', col='gray', lty=c(1,2,2), bty='l')
  abline(h=1)
  })
```
Observed vs expected
```{r}

predicted<-(sapply(inla1,'[[', 'posterior.pred.hdi', simplify=F) )
obs<- sapply(inla1,'[[', 'obs.y', simplify=F)
plot.func<-function(pred.ds, obs.ds){
    matplot(pred.ds ,type='l', col='gray', lty=c(1,2,2), bty='l')
  points(obs.ds, pch=16)
  abline(h=1)
}
mapply(plot.func, pred.ds=predicted, obs.ds=obs )
```



look at the PCs
```{r}
#library(devtools)
#install_github("vqv/ggbiplot")
#library(ggbiplot)
pcas<-lapply(inla1,'[[', 'pca.results')
lapply(pcas,ggbiplot )
#ggbiplot(pca1)
#screeplot(pca1)
```


